\plaintitle

### 2017-01-03 drafts/labor/cases.tex

{Research in crowdsourcing has spent the better part of a decade
exploring how to grow the limits of crowdsourcing,
finding the boundaries of crowd work and microtasks.}
This has largely involved iteratively
  identifying challenges to increasing complexity and
  overcoming them through novel designs of work--flows and processes~\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
The question that has emerged then
has been \textit{whether} there are limits to crowdsourcing
--- and if so, what factors determine them.
Through this lens,
a number of contributions to the field have pushed the boundaries of crowd work.

The exploration of crowdsourcing's potential and limits has principally looked at
manipulating and extending along three dimensions:
\begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
  \item \namerefl{sec:complexity}
  \item \namerefl{sec:decomposition}
  \item \namerefl{sec:relationships}.
\end{Numberlist}
We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
decomposed,  contextually abstracted, and
made more resilient to attrition of various forms.
We'll also point to corresponding piecework literature addressing these aspects.
\msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
Finally, we'll discuss how these elements will serve
to constrain the upper and lower bounds of crowdsourcing as it relates
to the question of the furthest limits of crowdsourcing.

@@

>>>>

{Research in crowdsourcing has spent the better part of a decade
exploring how to grow the limits of crowdsourcing,
finding the boundaries of crowd work and microtasks.}
This has largely involved iteratively
  identifying challenges to increasing complexity and
  overcoming them through novel designs of work--flows and processes (e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
The question that has emerged then
has been \textit{whether} there are limits to crowdsourcing
--- and if so, what factors determine them.
Through this lens,
a number of contributions to the field have pushed the boundaries of crowd work.

The exploration of crowdsourcing's potential and limits has principally looked at
manipulating and extending along three dimensions:
First, \namerefl{sec:complexity}.
Second, \namerefl{sec:decomposition}.
And third, \namerefl{sec:relationships}.
We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
decomposed,  contextually abstracted, and
made more resilient to attrition of various forms.
We'll also point to corresponding piecework literature addressing these aspects.
\msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
Finally, we'll discuss how these elements will serve
to constrain the upper and lower bounds of crowdsourcing as it relates
to the question of the furthest limits of crowdsourcing.

@@

\onlyinsubfile{
  \printbibliography{}
}

### 2017-01-03 drafts/labor/introduction.tex

\section{Introduction}\label{sec:introduction}
{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
constituent parts of larger tasks.}
This framing of work into modular blocks
has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in work whenever their schedules allow,
with little to no awareness of the broader context of the work, and
often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:
\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work
\item \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
\end{numberlist}
The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
like AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
\cite[e.g.][]{bernsteinSoylent,vizwiz,paolacci2010running}.

@@

\topic{This body of research has broadly sought to answer one central question:
What does the future hold for on--demand work and those who do it?}
Researchers have offered insights on this question along three major threads:
\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
  \item \nameref{sec:complexity}?
        Specifically,
        \begin{inlinelist}[label=(\alph*)]
          \item How complex are the goals that crowd work can accomplish?, and
          \item What kinds of goals and industries may eventually utilize it?
        \end{inlinelist}
        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
              Nebeling:2016:WCW:2858036.2858169,
              Hahn:2016:KAB:2858036.2858364}
  \item \nameref{sec:decomposition}
        \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
              LykourentzouPersonalityMatters,KucherbaevReLauncher,
              Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
  \item \nameref{sec:relationships}?
        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
              whyWouldAnyoneBrewer,takingAHITMcInnis}.
\end{numberlist}

@@

\subsection{Piecework as a lens to understand crowdsourcing}
In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework},
a work and payment structure which breaks tasks down into standalone contracts,
wherein payment is made for work output, rather than for time.
Piecework as a metaphor for crowd work is not new.
Indeed,
\citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
referenced crowd work as piecework briefly
as a loose analogy to the form of work emerging at the time~\cite{crowdworkFuture}.
But more than this,
the framing of on--demand labor as a re--instantiation of piecework
gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirically based research.

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by interrogating patterns that the corresponding literature predicts
on this basis, we can
\begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
  \item make sense of past events as part of a much larger series of an interrelated phenomenon
  \item reflect on differences in the factors that impacted piecework historically and their impact on--demand work today
  \item to some extent, offer predictions of what on--demand work researchers,
        and workers themselves,
        might expect to see on the horizon.
\end{numberlist}
For example, we will draw on the piecework literature regarding task decomposition,
which was historically limited by shortcomings in measurement and instrumentation, and
leverage that understanding to suggest how modern technology affects this mechanism in on--demand work.

We organize this paper as follows:
\begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
  \item we review the definition and historical arc of piecework
  to lay groundwork and make clear the analogy to on--demand work
  (which we'll call \textit{crowd work} hereon
  for consistency with the existing literature)
  \item we interrogate the three major research questions above using the lens of piecework. 
  We will identify similarities and differences between piecework as historically understood and
  on--demand work as we know it today
  \item we will make predictions of future developments based on how those similarities and differences influenced piecework
  \item we will offer implications for researchers and practitioners based on our results.
\end{numberlist}

>>>>

\section{Introduction}\label{sec:introduction}
\msb{Please take a full pass through for consistent terminology. In the intro, we align on on--demand work and call it the union of crowd work and gig work. But then, in the title and the research questions, everything is crowd work, crowd work, crowd work. I suggest either using crowd work in the introduction and explaining that you're using that to mean both traditional crowd work as well as modern gig work, or using on--demand work in the title and in the research questions. I see the point ``(which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature)'', but it feels like a cop out. Which one do we want to use, really? Let's pick one and be consistent throughout.}

\ali{\textit{On--demand work} is more ``correct'', but I'm worried that crowdsourcing researchers will skim it and feel there's nothing relevant to them. Is that giving them too little credit?}

\topic{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
constituent parts of larger tasks.}
This framing of work into modular
has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in work whenever their schedules allow,
with little to no awareness of the broader context of the work, and
often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:
first, \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
such as AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}).

@@

\topic{This body of research has broadly sought to answer one central question:
What does the future hold for on--demand work and those who do it?}
Researchers have offered insights on this question along three major threads:
% \begin{numberlist}[itemjoin*={.~And~},itemjoin={.~}]
  First, \nameref{sec:complexity}?
        Specifically,
        \begin{inlinelist}[label=(\alph*)]
          \item How complex are the goals that crowd work can accomplish?, and
          \item What kinds of goals and industries may eventually utilize it?
        \end{inlinelist}
        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
              Nebeling:2016:WCW:2858036.2858169,
              Hahn:2016:KAB:2858036.2858364}.
  Second, \nameref{sec:decomposition}
        \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
              LykourentzouPersonalityMatters,KucherbaevReLauncher,
              Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}.
  And third, \nameref{sec:relationships}?
        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
              whyWouldAnyoneBrewer,takingAHITMcInnis}.
% \end{numberlist}

@@

\subsection{Piecework as a lens to understand crowdsourcing}
In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework},
a work and payment structure which breaks tasks down into standalone contracts,
wherein payment is made for work output, rather than for time.
Piecework as a metaphor for crowd work is not new.
Indeed,
\citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
referenced crowd work as piecework briefly
as a loose analogy~\cite{crowdworkFuture}.
But more than this,
the framing of on--demand labor as a re--instantiation of piecework
gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirical research.

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by interrogating patterns that the corresponding literature predicts
on this basis, we can
% \begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
first, make sense of past events as part of a much larger series of an interrelated phenomenon;
second, reflect on differences in the factors that impacted piecework historically and their impact on--demand work today;
and third, to some extent, offer predictions of what on--demand work researchers,
and workers themselves,
might expect to see on the horizon.
% \end{numberlist}
For example, we will draw on the piecework literature regarding task decomposition,
which was historically limited by shortcomings in measurement and instrumentation, and
leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
--- namely, by enabling precise tracking and measurement of workers via algorithms and software.

We organize this paper as follows:
% \begin{numberlist}[itemjoin*={; and },itemjoin={;~}]
first, we review the definition and historical arc of piecework
to lay groundwork and make clear the analogy to on--demand work
(which we'll call \textit{crowd work} hereon
for consistency with the existing literature);
second, we interrogate the three major research questions above using the lens of piecework. 
We will identify similarities and differences between piecework as historically understood and
on--demand work as we know it today;
third, we will make predictions of future developments based on how those similarities and differences influenced piecework;
finally, we will offer implications for researchers and practitioners based on our results.
% \end{numberlist}

### 2017-01-03 drafts/labor/piecework_lit.tex

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{numberlist}[itemjoin*={, and },label={\nthwords*}]
  \item define ``piecework'' as researchers in its field understand it
  \item trace the rise and fall of piecework at a high level,
        identifying key figures and ideas during this time.
\end{numberlist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

% \onlyinsubfile{\clearpage}
\subsection{What is piecework?}
\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over time,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions of piecework,
tracing an outline of % a working understanding of
\textit{what piecework is}.

@@

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
\begin{Numberlist}
\item platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
\item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
\end{Numberlist}
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.

% \onlyinsubfile{\clearpage}
\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\topic{In this section we'll discuss the history of piecework at a high level
to contextualize the observations and conclusions we later make,
rather than provide a comprehensive historical account.}
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, it will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework;
as close to a \textit{thick description} of piecework
as we can afford~\cite{geertz1973interpretation}.

@@

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took off on in farm work~\cite{hughRaynbirdTaskWork},
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\citeauthor{hart2013rise} point out that the Second World War,
which called millions of Americans to military service,
necessitated the rapid training and employment of
a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.

The early growth of piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
Despite this, workers' needs were mostly ignored,
leading to frustration over poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}~\cite{hart2013rise}.

% \onlyinsubfile{\clearpage}
\subsection{Case studies in piecework}
We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
  \item railroad and other industrial workers across the United States
  \item Airy's employment of \textit{human computers}
  \item domestic and farm work (in particular, the ``match--girls'').
\end{numberlist}
We'll approach this by looking at familiar cases of decomposed, distributed work (industrial workers),
and make our way backwards through the \nth{20} and then \nth{19} century,
tracing the origins of the ideas that

@@

% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
% \onlyinsubfile{\clearpage}
\subsubsection{Railroad and other Industrial Workers}
{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
It was here that
scientific management, Fordism, and Taylorism
dramatically influenced how workers were managed and
the ways in which workers were perceived and envisioned paradigmatically;
it's here we'll start our overview of piecework's history through case studies.

\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

@@

\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.

% \onlyinsubfile{\clearpage}
\subsubsection{Domestic and Farmhand Labor}
\topic{When piecework entered the American economy, it was not used for complex work.}
One reason for low complexity was workers' skills:
it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted~\cite{hart2013rise}.
So, initially piecework in the United States arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
The complexity of the work remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
However, writers of the time focused their attention on
wage~\cite{burton1899commercial} and
management regimes~\cite{norton1900textile}
rather than training.
Mass manufacturing,
such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
flourished under piecework systems in densely populated cities.

\topic{Workers' relationships with employers quickly soured.}
The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described,
``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

% \onlyinsubfile{\clearpage}
\subsubsection{Airy's Computers}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
% \ali{also talk about the challenge/competition Airy was in with other astronomers,
% and the approach of verifying worker results}
% Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.
% For one thing,
Airy also instituted a policy of firing his computers once they reached age 23.

>>>>

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We will more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{numberlist}[itemjoin*={, and },label={\nthwords*}]
  \item define ``piecework'' as researchers in its field understand it
  \item trace the rise and fall of piecework at a high level,
        identifying key figures and ideas during this time.
\end{numberlist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

% \onlyinsubfile{\clearpage}
\subsection{What is piecework?}\label{sec:whatIsPiecework}
\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over time,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions of piecework,
tracing an outline of % a working understanding of
\textit{what piecework is}.

@@

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
\begin{Numberlist}
\item platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
\item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
\end{Numberlist}
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. 
\msb{This transition seems misplaced with the Piecework primer and Case Studies subsections up next:}
In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.

% \onlyinsubfile{\clearpage}
\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\msb{This section is so brief that it feels like a bit of a cheat. I think we need to frame it differently. Maybe that it should survey the different ``eras'' of piecework? And each paragraph is laying out a major historical era that we will return to later? Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
\ali{Do you mean that we'd have (sub)subsections under this? I'm tempted to merge this subsection into the ``\namerefl{sec:whatIsPiecework}'' subsection instead. Not exactly sure how that would work (i.e. would I interleave the history or just append it, basically removing the header).}
\topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
This section will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
% While the previous section provided a \textit{definition} of piecework,
% this section attempts to shine a light on the \textit{zeitgeist} of piecework;
% as close to a \textit{thick description} of piecework
% as we can afford~\cite{geertz1973interpretation}.
% \msb{is this really thick? as a reader it flies by really quickly and super high level...}

@@

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
\ali{I'm not sure; I'll look for some data.}
% \msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
% \citeauthor{hart2013rise} point out that the Second World War,
% which called millions of Americans to military service,
% necessitated the rapid training and employment of
% a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.

% \msb{The topic sentence is about management, but the content is about discontent. Which is it?}
\topic{The early growth of piecework led to discussion surrounding how best to manage pieceworkers,
generally in an adversarial approach, consequently
frustrating workers~\cite{norton1900textile,clark1908cotton}.}
This frustrations of poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

\msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\msb{How quickly did it plummet? By when was it gone? Where is it now?}
\msb{This sentence is empty calorie:}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
\msb{Very few calorie sentence:}
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance
\item the shifting of these industries (e.g., manufacturing, clothing)
to other countries
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}~\cite{hart2013rise}.

% \onlyinsubfile{\clearpage}
\subsection{Case studies in piecework}
Throughout the rest of the paper, we will return to three major case studies to frame our analyses:
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
first, railroad and other industrial workers;
second, Airy's employment of \textit{human computers};
and third, domestic and farm work (in particular, the ``match--girls'').
Here, we introduce the basic historical details of these three cases.
We begin with the most familiar case (industrial workers),
working backwards through the \nth{20} and then \nth{19} century,
tracing the origins of the ideas that \dots
% \msb{incomplete sentence}

@@

% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
% \onlyinsubfile{\clearpage}
\subsubsection{Railroad and other Industrial Workers}
{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
It was here that
scientific management, Fordism, and Taylorism
dramatically influenced how workers were managed and
the ways in which workers were perceived and envisioned paradigmatically;
it's here we'll start our overview of piecework's history through case studies. \msb{we're not starting an overview of history; aren't we midway through it?}

\msb{I'm feeling very disoriented in this paragraph. The subsection title is about railroads, but we're all fast-forwarded to auto manufacturing?}
\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

@@

\msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
\msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.

\msb{This subsection feels like it never delivered on what it promised: a case study of one group of pieceworkers. I thought the goal was to introduce the railroad workers --- who they were, what they did, what their conditions were like.}

% \onlyinsubfile{\clearpage}
\subsubsection{Domestic and Farmhand Labor}
\topic{When piecework entered the American economy, it was not used for complex work.}
One reason for low complexity was workers' skills:
it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted~\cite{hart2013rise}.
So, initially piecework in the United States arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
\msb{this is not about domestic/farmhand labor...might need to move up or be cut...or is it? clarify that this is domestic labor moving into NYC?}
The complexity of the work remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
\msb{how is this sentence relevant to the case study?:}
However, writers of the time focused their attention on
wage~\cite{burton1899commercial} and
management regimes~\cite{norton1900textile}
rather than training.
Mass manufacturing,
such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
flourished under piecework systems in densely populated cities.

\topic{Workers' relationships with employers quickly soured.}
\msb{We haven't met the match--girls yet. Is this case study supposed to be about them? They come out of nowhere, and we're talking about their strike rather than them.}
The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described,
``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
\msb{this paragraph seems to be about collective action rather than domestic/farmhand labor. also, coal miners aren't domestic or farmhand.}
Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
\msb{Why is Taylor relevant to this case study?}
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

% \onlyinsubfile{\clearpage}
\subsubsection{Airy's Computers}
\msb{This one seems way thinner than the others? Maybe it should be merged with the earlier description of Airy.}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
% \ali{also talk about the challenge/competition Airy was in with other astronomers,
% and the approach of verifying worker results}
% Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.
% For one thing,
Airy also instituted a policy of firing his computers once they reached age 23.
